---
title: "R Notebook"
output: html_notebook
---
Importing Coffee chains data and
Descriptive Statistics

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Importing the data

```{r}
setwd("~/Documents/class/stats-programming/week-7")
data <- read.csv("dataset.csv") %>% clean_names() %>%
  mutate(mode = str_replace(as.character(mode), "0", "minor"),
         mode = str_replace(as.character(mode), "1", "major"),
         time_signature = as.character(time_signature) ) %>% 
  select(-1)

str_replace

         
dataset<-data
dd <- data

names(data)

#data_full <- data %>% mutate(
                #cogs_proportion = cogs/sales,
                #inventory_proportion = inventory/sales,
                #margin_proportion = margin/sales,
                #marketing_proportion = marketing/sales,
                #profit_proportion = profit/sales,
                #expense_proportion = total_expenses/sales,
                #area_code = as.character(area_code),
                #date = as.Date(date, origin = "1899-12-30")
               # ) %>% filter(date == "2013-12-01") %>% 
  #select(1:9, 14:20)
    
```

Note, the data covers monthly reports from Jan 2012 to Dec 2013.
In order to do PCA, maybe we can filter to just one month's worth of data.

```{r}
#data_for_pca_analysis <- data_full %>%
# select(1:9, 19, 21:26) %>%
 # filter(date == "2013-12-01")

#data_for_pca_analysis
```

```{r}

dim(dd)
n<-dim(dd)[1]
K<-dim(dd)[2]

descriptiva<-function(X, nom){
  if (!(is.numeric(X) || class(X)=="Date")){ 
    frecs<-table(as.factor(X), useNA="ifany")
    proportions<-frecs/n
    #ojo, decidir si calcular porcentages con o sin missing values
    pie(frecs, cex=0.6, main=paste("Pie of", nom))
    barplot(frecs, las=3, cex.names=0.7, main=paste("Barplot of", nom), col=listOfColors)
    print(paste("Number of modalities: ", length(frecs)))
    print("Frequency table")
    print(frecs)
    print("Relative frequency table (proportions)")
    print(proportions)
    print("Frequency table sorted")
    print(sort(frecs, decreasing=TRUE))
    print("Relative frequency table (proportions) sorted")
    print(sort(proportions, decreasing=TRUE))
   }else{
     if(class(X)=="Date"){
       print(summary(X))
       print(sd(X))
       #decide breaks: weeks, months, quarters...
       hist(X,breaks="weeks")
     }else{
       hist(X, main=paste("Histogram of", nom))
       boxplot(X, horizontal=TRUE, main=paste("Boxplot of",nom))
       print("Extended Summary Statistics")
       print(summary(X))
       print(paste("sd: ", sd(X, na.rm=TRUE)))
       print(paste("vc: ", sd(X, na.rm=TRUE)/mean(X, na.rm=TRUE)))
      }
   }
}

actives<-c(1:K)
colDate<-1
#if (dataset=="platjaDaro")
#  {dd[,colDate]<-as.Date(dd[, colDate], format="%d/%m/%y %h:%m:%s")
#   actives<-c(3:44)
#}

listOfColors<-rainbow(39)

par(ask=TRUE)

for(k in actives){
  print(paste("variable ", k, ":", names(dd)[k] ))
  descriptiva(dd[,k], names(dd)[k])
}
par(ask=FALSE)
```

# Principal component Analysis

```{r}
objects()
attributes(dd)

attach(dd)
names(dd)

sapply(dd,class)

numerical<-which(sapply(dd,is.numeric))
numerical

dcon<-dd[,numerical]
sapply(dcon,class)

# PRINCIPAL COMPONENT ANALYSIS OF dcon

pc1 <- prcomp(dcon, scale=TRUE, center = TRUE)
class(pc1)
attributes(pc1)

print(pc1)

pc1

str(pc1)

```

```{r}
plot(pc1, main = "") 
```


```{r}
# WHICH PERCENTAGE OF THE TOTAL INERTIA IS REPRESENTED IN SUBSPACES?

pc1$sdev
inerProj<- pc1$sdev^2 
inerProj
totalIner<- sum(inerProj)
totalIner
pinerEix<- 100*inerProj/totalIner
pinerEix
barplot(pinerEix)

#Cummulated Inertia in subspaces, from first principal component to the 11th dimension subspace
barplot(100*cumsum(pc1$sdev[1:dim(dcon)[2]]^2)/dim(dcon)[2])
percInerAccum<-100*cumsum(pc1$sdev[1:dim(dcon)[2]]^2)/dim(dcon)[2]
percInerAccum
```

```{r}
# SELECTION OF THE SINGIFICNT DIMENSIONS (keep 80% of total inertia)

nd = 10

print(pc1)
attributes(pc1)
pc1$rotation

# STORAGE OF THE EIGENVALUES, EIGENVECTORS AND PROJECTIONS IN THE nd DIMENSIONS
View(pc1$x)
dim(pc1$x)
dim(dcon)
dcon[100,]
pc1$x[100,]

Psi = pc1$x[,1:nd]
dim(Psi)
Psi[100,]

# STORAGE OF LABELS FOR INDIVIDUALS AND VARIABLES

iden = row.names(dcon)
etiq = names(dcon) # getting names of numerical variables
ze = rep(0,length(etiq)) # WE WILL NEED THIS VECTOR AFTERWARDS FOR THE GRAPHICS

names(dcon)
# PLOT OF INDIVIDUALS

#select your axis
#eje1<-2
eje1<-1
#eje2<-3
eje2<-2


plot(Psi[,eje1],Psi[,eje2])
text(Psi[,eje1],Psi[,eje2],labels=iden, cex=0.5)
axis(side=1, pos= 0, labels = F, col="cyan")
axis(side=3, pos= 0, labels = F, col="cyan")
axis(side=2, pos= 0, labels = F, col="cyan")
axis(side=4, pos= 0, labels = F, col="cyan")


plot(Psi[,eje1],Psi[,eje2], type="n")
text(Psi[,eje1],Psi[,eje2],labels=iden, cex=0.5)
axis(side=1, pos= 0, labels = F, col="cyan")
axis(side=3, pos= 0, labels = F, col="cyan")
axis(side=2, pos= 0, labels = F, col="cyan")
axis(side=4, pos= 0, labels = F, col="cyan")
#library(rgl)
#plot3d(Psi[,1],Psi[,2],Psi[,3])
```

```{r}

#Projection of variables

Phi = cor(dcon,Psi)
Phi

X<-Phi[,eje1]
Y<-Phi[,eje2]

plot(Psi[,eje1],Psi[,eje2],type="n")
axis(side=1, pos= 0, labels = F)
axis(side=3, pos= 0, labels = F)
axis(side=2, pos= 0, labels = F)
axis(side=4, pos= 0, labels = F)
arrows(ze, ze, X, Y, length = 0.07,col="blue")
text(X,Y,labels=etiq,col="darkblue", cex=0.7)

#zooms
plot(Psi[,eje1],Psi[,eje2],type="n",xlim=c(min(X,0),max(X,0)), ylim=c(-1,1))
axis(side=1, pos= 0, labels = F)
axis(side=3, pos= 0, labels = F)
axis(side=2, pos= 0, labels = F)
axis(side=4, pos= 0, labels = F)
arrows(ze, ze, X, Y, length = 0.07,col="blue")
text(X,Y,labels=etiq,col="darkblue", cex=0.7)


```
Trying to plot based on genre

```{r}
varcat=factor(dd[,21])
plot(Psi[,1],Psi[,2],col=varcat)
axis(side=1, pos= 0, labels = F, col="darkgray")
axis(side=3, pos= 0, labels = F, col="darkgray")
axis(side=2, pos= 0, labels = F, col="darkgray")
axis(side=4, pos= 0, labels = F, col="darkgray")
legend("bottomleft",levels(factor(varcat)),pch=1,col=c(1:114), cex=0.6)
```

Trying to plot based on explicitness

```{r}
varcat=factor(dd[,8])
plot(Psi[,1],Psi[,2],col=varcat)
axis(side=1, pos= 0, labels = F, col="darkgray")
axis(side=3, pos= 0, labels = F, col="darkgray")
axis(side=2, pos= 0, labels = F, col="darkgray")
axis(side=4, pos= 0, labels = F, col="darkgray")
legend("bottomleft",levels(factor(varcat)),pch=1,col=c(1,2,3,4), cex=0.6)
```

Plotting based on major/minor
```{r}
varcat=factor(dd[,12])
plot(Psi[,1],Psi[,2],col=varcat)
axis(side=1, pos= 0, labels = F, col="darkgray")
axis(side=3, pos= 0, labels = F, col="darkgray")
axis(side=2, pos= 0, labels = F, col="darkgray")
axis(side=4, pos= 0, labels = F, col="darkgray")
legend("bottomleft",levels(factor(varcat)),pch=1,col=c(1,2,3,4), cex=0.6)
```


```{r}
plot(Psi[,eje1],Psi[,eje2],type="n",xlim=c(-1,1), ylim=c(-3,1))
#plot(X,Y,type="none",xlim=c(min(X,0),max(X,0)))
axis(side=1, pos= 0, labels = F, col="cyan")
axis(side=3, pos= 0, labels = F, col="cyan")
axis(side=2, pos= 0, labels = F, col="cyan")
axis(side=4, pos= 0, labels = F, col="cyan")

arrows(ze, ze, X, Y, length = 0.07,col="lightgray")
text(X,Y,labels=etiq,col="gray", cex=0.7)

#all qualitative together
plot(Psi[,eje1],Psi[,eje2],type="n")
axis(side=1, pos= 0, labels = F, col="cyan")
axis(side=3, pos= 0, labels = F, col="cyan")
axis(side=2, pos= 0, labels = F, col="cyan")
axis(side=4, pos= 0, labels = F, col="cyan")

#nominal qualitative variables

dcat<-c(3, 4, 5, 6, 7, 8, 9)

c<-1
for(k in dcat){
  seguentColor<-colors[c]
  
  fdic1 = tapply(Psi[,eje1],dd[,k],mean)
  fdic2 = tapply(Psi[,eje2],dd[,k],mean) 
  
  points(fdic1,fdic2,pch=16,col=seguentColor, labels=levels(dd[,k]))
  text(fdic1,fdic2,col=seguentColor, cex=0.6)
  c<-c+1
}
legend("bottomleft",names(dd)[dcat],pch=1,col=colors, cex=0.6)


dordi<-c(8)


levels(factor(dd[,dordi[1]]))
#reorder modalities: when required
dd[,dordi[1]] <- factor(dd[,dordi[1]], ordered=TRUE,  levels= c("WorkingTypeUnknown","altres sit","temporal","fixe","autonom"))
levels(dd[,dordi[1]])

c<-1
col<-1
for(k in dordi){
  seguentColor<-colors[col]
  fdic1 = tapply(Psi[,eje1],dd[,k],mean)
  fdic2 = tapply(Psi[,eje2],dd[,k],mean) 
  
  #points(fdic1,fdic2,pch=16,col=seguentColor, labels=levels(dd[,k]))
  #connect modalities of qualitative variables
  lines(fdic1,fdic2,pch=16,col=seguentColor)
 text(fdic1,fdic2,labels=levels(dd[,k]),col=seguentColor, cex=0.6)
  c<-c+1
  col<-col+1
}
legend("topleft",names(dd)[dordi],pch=1,col=colors[1:length(dordi)], cex=0.6)

```
The following code is to identify clusters in the pca

```{r}

colors<-c("red", "blue", "darkgreen", "orange", "violet", "magenta", "pink")

#represent numerical variables in background
plot(Psi[,eje1],Psi[,eje2],type="n",xlim=c(-1,1), ylim=c(-3,1))
#plot(X,Y,type="none",xlim=c(min(X,0),max(X,0)))
axis(side=1, pos= 0, labels = F, col="cyan")
axis(side=3, pos= 0, labels = F, col="cyan")
axis(side=2, pos= 0, labels = F, col="cyan")
axis(side=4, pos= 0, labels = F, col="cyan")

#add projections of numerical variables in background
arrows(ze, ze, X, Y, length = 0.07,col="lightgray")
text(X,Y,labels=etiq,col="gray", cex=0.7)

#add centroids
c<-1
for(k in dcat){
  seguentColor<-colors[c]
  
  fdic1 = tapply(Psi[,eje1],dd[,k],mean)
  fdic2 = tapply(Psi[,eje2],dd[,k],mean) 
  
  #points(fdic1,fdic2,pch=16,col=seguentColor, labels=levels(dd[,k]))
  text(fdic1,fdic2,col=seguentColor, cex=0.6)
  c<-c+1
}
legend("bottomleft",names(dd)[dcat],pch=19,col=colors, cex=0.6)


#add ordinal qualitative variables. Ensure ordering is the correct

dordi<-c(8)


levels(factor(dd[,dordi[1]]))
#reorder modalities: when required
dd[,dordi[1]] <- factor(dd[,dordi[1]], ordered=TRUE,  levels= c("WorkingTypeUnknown","altres sit","temporal","fixe","autonom"))
levels(dd[,dordi[1]])

c<-1
col<-length(dcat)+1
for(k in dordi){
  seguentColor<-colors[col]
  fdic1 = tapply(Psi[,eje1],dd[,k],mean)
  fdic2 = tapply(Psi[,eje2],dd[,k],mean) 
  
  #points(fdic1,fdic2,pch=16,col=seguentColor, labels=levels(dd[,k]))
  #connect modalities of qualitative variables 
  lines(fdic1,fdic2,pch=16,col=seguentColor)
  text(fdic1,fdic2,labels=levels(dd[,k]),col=seguentColor, cex=0.6)
  c<-c+1
  col<-col+1
}
legend("topleft",names(dd)[dordi],pch=19,col=colors[col:col+length(dordi)-1], cex=0.6)


#Make two complementary factorial maps

colors<-c("red", "blue", "darkgreen", "orange", "violet", "magenta", "pink")

#represent numerical variables in background
#plot(Psi[,eje1],Psi[,eje2],type="p",xlim=c(-1,1), ylim=c(-3,1), col="lightgray")
plot(Psi[,eje1],Psi[,eje2],type="n",xlim=c(-1,1), ylim=c(-3,1))
#plot(X,Y,type="none",xlim=c(min(X,0),max(X,0)))
axis(side=1, pos= 0, labels = F, col="cyan")
axis(side=3, pos= 0, labels = F, col="cyan")
axis(side=2, pos= 0, labels = F, col="cyan")
axis(side=4, pos= 0, labels = F, col="cyan")
R...